---
title: fires-2020-05-07
author: 
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(sf)
library(lwgeom)

# Parameters

# FRAP
# CALFIRE perimeters
# Source: https://frap.fire.ca.gov/frap-projects/fire-perimeters/
file_frap <- here::here("data-raw", "fire19_1.gdb")

# Directory of tract shapefiles
dir_nhgis <- here::here("data-raw", "nhgis0001_shape")

# SIT-209
file_sit209 <- here::here("data-raw", "wildfires_deduped.csv")

#===============================================================================

# Code

# Read in the FRAP data
frap <-
  st_read(dsn = file_frap, layer = "firep19_1") %>% 
  as_data_frame() %>% 
  st_as_sf() 

# List of files in the NHGIS directory
dir_list <- fs::dir_ls(dir_nhgis)

# Read in each of the files in the NHGIS directory
tracts_2000 <- read_sf(dir_list[1])
tracts_2010 <- read_sf(dir_list[2])
tracts_2011 <- read_sf(dir_list[3])
tracts_2012 <- read_sf(dir_list[4])
tracts_2013 <- read_sf(dir_list[5])
tracts_2014 <- read_sf(dir_list[6])
tracts_2015 <- read_sf(dir_list[7])
tracts_2016 <- read_sf(dir_list[8])
tracts_2017 <- read_sf(dir_list[9])
# This doesn't read in; I have to diagnose why
# TODO: Find different boundaries
# tracts_2018 <- read_sf(dir_list[10])

# Read in the SIT-209 data
sit209 <-
  file_sit209 %>% 
  read_csv(
    col_types = 
      cols(
        .default = col_character(),
        LATITUDE = col_double(),
        LONGITUDE = col_double(),
        #START_DATE = col_datetime(format = "%m/%d/%y 00:00:00"),
        ESTIMATED_COST = col_double(),
        REPORT_DATE = col_datetime(format = ""),
        MAX_PERSONNEL = col_double(),
        INCIDENT_AREA = col_double(),
        MAX_SFR_THREAT = col_double(),
        MAX_MFR_THREAT = col_double(),
        MAX_MU_THREAT = col_double(),
        MAX_COMM_THREAT = col_double(),
        MAX_OUTB_THREAT = col_double(),
        MAX_OTHER_THREAT = col_double(),
        SFR_DAMAGED = col_double(),
        MFR_DAMAGED = col_double(),
        MU_DAMAGED = col_double(),
        COMM_DAMAGED = col_double(),
        OTHER_DAMAGED = col_double(),
        SFR_DESTROYED = col_double(),
        MFR_DESTROYED = col_double(),
        MU_DESTROYED = col_double(),
        COMM_DESTROYED = col_double(),
        OUTB_DESTROYED = col_double(),
        OTHER_DESTROYED = col_double(),
        INJURIES = col_double(),
        FATALITIES = col_double(),
        REPORT_ID = col_character(),
        WILDFIRE = col_logical(),
        STATE_FIPS = col_character(),
        DESTROYED = col_logical(),
        THREATENED = col_logical(),
        CLIMATE_REGION = col_character()
      )
  )
```

### Join the tract-level data together

Now that we've read in all of the tract data, let's join the files into one tibble.

```{r}
# tracts <-
#   rbind(
#     tracts_2000 %>% mutate(year = 2000) %>% select(year, GEOID = CTIDFP00),
#     tracts_2010 %>% mutate(year = 2010) %>% select(year, GEOID = GEOID10),
#     tracts_2011 %>% mutate(year = 2011) %>% select(year, GEOID),
#     tracts_2012 %>% mutate(year = 2012) %>% select(year, GEOID),
#     tracts_2013 %>% mutate(year = 2013) %>% select(year, GEOID),
#     tracts_2014 %>% mutate(year = 2014) %>% select(year, GEOID),
#     tracts_2015 %>% mutate(year = 2015) %>% select(year, GEOID),
#     tracts_2016 %>% mutate(year = 2016) %>% select(year, GEOID),
#     tracts_2017 %>% mutate(year = 2017) %>% select(year, GEOID)
#   ) %>% 
#   filter(str_starts(GEOID, "06")) # Limits us to tracts in California.
```

As an optional step, we'll write out this tibble for future use.

```{r}
# tracts %>% write_sf(path = here::here("data", "ca_tracts_2000_2017"))
```

### FRAP

First, I prepare the FRAP data to be joined to the Census tract data:

* I convert `YEAR_` to an integer from a factor.
* After a quick check, it looks like neither FIRE_NUM, nor INC_NUM, nor FIRE_NAME in a given year are unique, so I make our own unique id.
* I fiddle with the geography column a little to avoid errors.
* I toss out duplicate values

```{r}
frap_prepped <-
  frap %>% 
  mutate(
    YEAR_ = as.integer(as.character(YEAR_)),
    ALARM_DATE = 
      if_else(is.na(ALARM_DATE), "no_date", as.character(ALARM_DATE)),
    FIRE_NAME = 
      if_else(is.na(FIRE_NAME), "no_name", as.character(FIRE_NAME)),
    INC_NUM = 
      if_else(is.na(INC_NUM), "no_inc_num", as.character(INC_NUM)),
    GIS_ACRES = 
      if_else(is.na(GIS_ACRES), "no_acres", as.character(GIS_ACRES)),
    unique_id = 
      str_c(
        ALARM_DATE,
        FIRE_NAME,
        INC_NUM,
        GIS_ACRES
      )
  ) %>% 
  st_cast("MULTIPOLYGON") %>% # Cast to the same type as the tracts
  st_make_valid() %>% 
  distinct(unique_id, .keep_all = TRUE)
```

Let's try to join just 2000 census tract boundaries to the frap data.

```{r}
frap_2000 <-
  frap_prepped %>% 
  filter(
    YEAR_ <= 2009,
    YEAR_ >= 2000
  )

frap_tracts_2000 <-
  tracts_2000 %>% 
  filter(STATEFP00 == "06") %>% # Limit to California Census tracts
  select(GEOID = CTIDFP00) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(frap_2000)
```

Check that no duplicates were created:

```{r}
frap_tracts_2000 %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

We combine 2010:

```{r}
frap_2010 <-
  frap_prepped %>% 
  filter(YEAR_ == 2010)

frap_tracts_2010 <-
  tracts_2010 %>% 
  filter(STATEFP10 == "06") %>% # Limit to California Census tracts
  select(GEOID = GEOID10) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(frap_2010)
```

Check for problems: 

```{r}
frap_tracts_2010 %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

We combine 2011:

```{r}
frap_2011 <-
  frap_prepped %>% 
  filter(YEAR_ == 2011)

frap_tracts_2011 <-
  tracts_2011 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(frap_2011)
```

Check for problems: 

```{r}
frap_tracts_2011 %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

We combine 2012:

```{r}
frap_2012 <-
  frap_prepped %>% 
  filter(YEAR_ == 2012)

frap_tracts_2012 <-
  tracts_2012 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(frap_2012)
```

Check for problems: 

```{r}
frap_tracts_2012 %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

We combine 2013:

```{r}
frap_2013 <-
  frap_prepped %>% 
  filter(YEAR_ == 2013)

frap_tracts_2013 <-
  tracts_2013 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(frap_2013)
```

Check for problems: 

```{r}
frap_tracts_2013 %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

We combine 2014:

```{r}
frap_2014 <-
  frap_prepped %>% 
  filter(YEAR_ == 2014)

frap_tracts_2014 <-
  tracts_2014 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(frap_2014)
```

Check for problems: 

```{r}
frap_tracts_2014 %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

We combine 2015:

```{r}
frap_2015 <-
  frap_prepped %>% 
  filter(YEAR_ == 2015)

frap_tracts_2015 <-
  tracts_2015 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(frap_2015)
```

Check for problems: 

```{r}
frap_tracts_2015 %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

We combine 2016:

```{r}
frap_2016 <-
  frap_prepped %>% 
  filter(YEAR_ == 2016)

frap_tracts_2016 <-
  tracts_2016 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(frap_2016)
```

Check for problems: 

```{r}
frap_tracts_2016 %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

We combine 2017:

```{r}
frap_2017 <-
  frap_prepped %>% 
  filter(YEAR_ == 2017)

frap_tracts_2017 <-
  tracts_2017 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(frap_2017)
```

Check for problems: 

```{r}
frap_tracts_2017 %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

Put it all together!

```{r}
frap_tracts <-
  rbind(
    frap_tracts_2000,
    frap_tracts_2010,
    frap_tracts_2011,
    frap_tracts_2012,
    frap_tracts_2013,
    frap_tracts_2014,
    frap_tracts_2015,
    frap_tracts_2016,
    frap_tracts_2017
  )
```

Check for duplicates:

```{r}
frap_tracts %>% 
  st_drop_geometry() %>% 
  count(unique_id, GEOID, sort = TRUE) %>% 
  filter(n > 1)
```

Get rid of the duplicates:

```{r}
frap_tracts_distinct <-
  frap_tracts %>% 
  distinct(unique_id, .keep_all = TRUE)
  
```

Write out the final product:

```{r}
frap_tracts_distinct %>%
  st_drop_geometry() %>% 
  write_csv(here::here("data", "frap_tracts.csv"))
```

### SIT-209

First, I prepare the SIT-209, filtering to just the data we need:

* Wildfire incidents only
* Wildfires in which at least one structure was destroyed
* Only in California

Then I take steps to help us to join this data with other datasets:

* I transform latitude and longitude to correct any points that need to be standardized.
* I make sure the COUNTY and STATE_FIPS columns are standardized to correspond to the frap_tracts_distinct dataframe.

```{r}
sit209_prepped <-
  sit209 %>% 
  filter(
    WILDFIRE == TRUE,
    DESTROYED == TRUE,
    STATE == "CA"
  ) %>% 
  mutate( # Standardized latitude and longitude into the correct hemisphere
    LATITUDE = if_else(LATITUDE < 0, LATITUDE * -1, LATITUDE),
    LONGITUDE = if_else(LONGITUDE > 0, LONGITUDE * -1, LONGITUDE),
    COUNTY = str_to_title(COUNTY),
    STATE_FIPS = map(STATE_FIPS, ~ if_else(length(.) < 2, str_c("0", .),.)),
    STATE_FIPS = pluck(STATE_FIPS, 1)
  ) 
```

But here's the problem: There are often fires in the same county and month, so we couldn't join on these criteria alone.

```{r}
sit209_prepped %>%
  mutate(
    report_month = lubridate::month(REPORT_DATE),
  ) %>% 
  count(COUNTY, REPORT_YEAR, report_month, sort = TRUE)
```

This is true in the FRAP data, too.

```{r}
frap_tracts_distinct %>%
  st_drop_geometry() %>% 
  mutate(
    ALARM_DATE = str_remove(ALARM_DATE, " 00:00:00"),
    ALARM_DATE = parse_date(ALARM_DATE, format = "%Y-%m-%d"),
    alarm_month = lubridate::month(ALARM_DATE),
    county = str_sub(GEOID, 1L, 5L)
  ) %>% 
  count(county, alarm_month, YEAR_)
```

Counting units does not work any better:

```{r}
sit209_prepped %>% 
  mutate(
    report_month = lubridate::month(REPORT_DATE),
    unit_id = str_extract(INCIDENT_NUMBER, "-\\w{3}-")
  ) %>%
  count(unit_id, REPORT_YEAR, report_month, sort = TRUE)
```

```{r}
frap_tracts_distinct %>% 
  st_drop_geometry() %>% 
  mutate(
    ALARM_DATE = str_remove(ALARM_DATE, " 00:00:00"),
    ALARM_DATE = parse_date(ALARM_DATE, format = "%Y-%m-%d"),
    alarm_month = lubridate::month(ALARM_DATE),
    county = str_sub(GEOID, 1L, 5L)
  ) %>% 
  count(UNIT_ID, alarm_month, YEAR_, sort = TRUE)
```

So, maybe latitude and longitude are the ticket.

First, let's check out how many of the points of origin given in the latitude and longitude are outside of California:

```{r}
sit209_prepped %>% 
  # Use lat/long to make an sf object for spatial operations
  drop_na(LONGITUDE, LATITUDE) %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4269) %>% 
  st_difference(
    ussf::boundaries(geography = c("state"), projection = c("longlat")) %>% 
    st_transform(crs = 4269) %>% 
    filter(STATEFP == "06")
  ) %>% 
  count(INCIDENT_NAME, INCIDENT_NUMBER, sort = TRUE)
```

Looks like there are 27 fires with the wrong latitude and longitude. 

Then, there are 11 incidents where we don't have latitude or longitude:

```{r}
sit209_prepped %>% 
  count(is.na(LATITUDE) | is.na(LONGITUDE))
```

So, in total, we'll have to place 48 fires manually. Let's filter out these misfits and then join the rest.

```{r}
wrong_points <-
  sit209_prepped %>% 
  # Use lat/long to make an sf object for spatial operations
  drop_na(LONGITUDE, LATITUDE) %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4269) %>% 
  st_difference(
    ussf::boundaries(geography = c("state"), projection = c("longlat")) %>% 
    st_transform(crs = 4269) %>% 
    filter(STATEFP == "06")
  ) %>% 
  pull(INCIDENT_NUMBER)

```

* Join each tract year to sit209
* Join SIT209 with tracts to FRAP on tracts

```{r}
sit209_points <-
  sit209_prepped %>% 
  filter(!INCIDENT_NUMBER %in% wrong_points) %>% 
  drop_na(LONGITUDE, LATITUDE) %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 3310) %>% 
  st_transform(crs = 3310)
```


```{r}
sit209_tracts_2000 <-
  tracts_2000 %>% 
  filter(STATEFP00 == "06") %>% # Limit to California Census tracts
  select(GEOID = CTIDFP00) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(sit209_points %>% filter(REPORT_YEAR <= 2000))
```


Let's do the rest:

```{r}
sit209_tracts_2010 <-
  tracts_2010 %>% 
  filter(STATEFP10 == "06") %>% # Limit to California Census tracts
  select(GEOID = GEOID10) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(sit209_points %>% filter(REPORT_YEAR == 2010))

sit209_tracts_2011 <-
  tracts_2011 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(sit209_points %>% filter(REPORT_YEAR == 2011))

sit209_tracts_2012 <-
  tracts_2012 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>%  
  st_intersection(sit209_points %>% filter(REPORT_YEAR == 2012))

sit209_tracts_2013 <-
  tracts_2013 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(sit209_points %>% filter(REPORT_YEAR == 2013))

sit209_tracts_2014 <-
  tracts_2014 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(sit209_points %>% filter(REPORT_YEAR == 2014))

sit209_tracts_2015 <-
  tracts_2015 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(sit209_points %>% filter(REPORT_YEAR == 2015))

sit209_tracts_2016 <-
  tracts_2016 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(sit209_points %>% filter(REPORT_YEAR == 2016))

sit209_tracts_2017 <-
  tracts_2017 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_intersection(sit209_points %>% filter(REPORT_YEAR == 2017))

```

NOW WE PUT IT ALL TOGETHER.

```{r}
sit209_tracts <-
  rbind(
    sit209_tracts_2000,
    sit209_tracts_2010,
    sit209_tracts_2011,
    sit209_tracts_2012,
    sit209_tracts_2013,
    sit209_tracts_2014,
    sit209_tracts_2015,
    sit209_tracts_2016,
    sit209_tracts_2017
  )
```

The problem is that this tibble excludes so many points, and I'm not sure why. I guess they're just not in the tracts? That doesn't seem true. Something funny is happening.

Adding a buffer does not help:

```{r}
sit209_points %>% 
  filter(REPORT_YEAR == 2017) %>% 
  st_buffer(100) %>% 
  st_intersection(
    tracts_2017 %>% 
    filter(STATEFP == "06") %>% # Limit to California Census tracts
    select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
    st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
    st_make_valid()
  )
```

Maybe it's the intersection that isn't working. Here's st_within:

```{r}
sit209_points %>% 
  filter(REPORT_YEAR == 2017) %>% 
  st_within(
    tracts_2017 %>% 
    filter(STATEFP == "06") %>% # Limit to California Census tracts
    select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
    st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
    st_make_valid()
  )
```

This did not work. How about st_contains?

```{r}
tracts_2017 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  st_join(sit209_points %>% filter(REPORT_YEAR == 2017), join = st_contains) %>% 
  drop_na(REPORT_YEAR)
```

Same darn result.

How about st_is_within_distance?

```{r}
tracts_2017 %>% 
  filter(STATEFP == "06") %>% # Limit to California Census tracts
  select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
  st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
  st_make_valid() %>% 
  mutate(centroid = st_centroid(geometry)) %>% 
  st_join(
    sit209_points %>% filter(REPORT_YEAR == 2017), 
    join = st_is_within_distance, dist = 100
  ) %>% 
  drop_na(REPORT_YEAR)

```

This didn't work either. Let's try a slightly different test:

```{r}

st_is_within_distance(
  x = 
    tracts_2017 %>% 
      filter(STATEFP == "06") %>% # Limit to California Census tracts
      select(GEOID) %>% # Selects and renames only the GEOID and the geometry column
      st_transform(crs = 3310) %>% # Makes it a California Albers, like the FRAP data.
      st_make_valid() %>% 
      mutate(centroid = st_centroid(geometry)),
  y = sit209_points %>% filter(REPORT_YEAR == 2017),
  dist = 100
) %>%
as_tibble()   

```

This also did not work.

```{r}
frap_tracts_distinct %>% 
  st_drop_geometry() %>%
  mutate(
    YEAR_ = as.character(YEAR_),
    ALARM_DATE = str_remove(ALARM_DATE, " 00:00:00"),
    ALARM_DATE = parse_date(ALARM_DATE, format = "%Y-%m-%d"),
    alarm_month = lubridate::month(ALARM_DATE)
  ) %>% 
  left_join(
    sit209_tracts %>% 
      st_drop_geometry() %>% 
      mutate(report_month = lubridate::month(REPORT_DATE)),
    by = c("GEOID", "YEAR_" = "REPORT_YEAR", "alarm_month" = "report_month")
  ) %>% 
  drop_na(DESTROYED)
```

This outcome is really disappointing. Let's try something else.
