---
title: 
author: 
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(sf)
library(tidycensus)
library(dbplyr)

# Parameters

## All historical perimeters
file_geomac <- here::here("data-raw", "US_HIST_FIRE_PERIMTRS_DD83", "US_HIST_FIRE_PERIMTRS_DD83.shp")
## 2000 SVI
file_svi_2000 <- here::here("data-raw", "SVI2000_US.csv")
## 2010 SVI
file_svi_2010 <- here::here("data-raw", "SVI2010_US.csv")
## 2014 SVI
file_svi_2014 <- here::here("data-raw", "SVI2014_US", "SVI2014_US.shp")
## 2016 SVI
file_svi_2016 <- here::here("data-raw", "SVI2016_US", "SVI2016_US.shp")
## 2018 SVI
file_svi_2018 <- here::here("data-raw", "SVI2018_US", "SVI2018_US_tract.shp")
# SIT-209: 2014 - 2017
# Source: BigLocalNews
# Download date: 2020-04-22
# Years: 2014 - 2017
url_sit209 <- "https://stacks.stanford.edu/file/druid:xj043rd8767/xj043rd8767_fire_reports_2014-2017.csv"
# Monitoring Trends in Burn Severity: Burned Areas Boundary Dataset
# Source: https://www.mtbs.gov/direct-download
# Download date: 2020-04-21
# Geography: United States
# Years: 2006 - 2017
file_fires <- here::here("data-raw", "mtbs_perimeter_data", "mtbs_perims_DD.shp")


#===============================================================================

# Code

# Read in the MTBS data
fires <-
  read_sf(file_fires)

## Read in the GEOMAC data, slice it into years

geomac <-
  read_sf(file_geomac)

geomac_2000 <-
  geomac %>% 
  filter(year_ == "2000")

geomac_2010 <-
  geomac %>% 
  filter(year_ == "2010")

geomac_2014 <-
  geomac %>% 
  filter(year_ == "2014")

geomac_2016 <-
  geomac %>% 
  filter(year_ == "2016")

geomac_2018 <-
  geomac %>% 
  filter(year_ == "2018")

## Read in the SVI data

svi_2000 <-
  file_svi_2000 %>% 
  read_csv()

svi_2010 <-
  file_svi_2010 %>% 
  read_csv()

svi_2014 <-
  read_sf(file_svi_2014)

svi_2016 <-
  read_sf(file_svi_2016)

svi_2018 <-
  read_sf(file_svi_2018)

## Get tract boundaries to join with 2000, 2010 csvs
## This is an extra step required for these two years only

# us <- unique(fips_codes$state)[1:51]
# 
# totalpop <- 
#   map_df(
#     us, 
#     function(x) {
#       get_acs(geography = "tract", variables = "B01003_001", state = x, geometry = TRUE)
#     }
#   )
# 
# svi_2000_geo <-
#   svi_2000 %>% 
#   left_join(totalpop, by = c("FIPS" = "GEOID")) %>% 
#   st_as_sf() 
# 
# svi_2010_geo <-
#   svi_2010 %>% 
#   left_join(totalpop, by = c("FIPS" = "GEOID")) %>% 
#   st_as_sf()

## Buffer and join the 2000 data

# svi_2000_buff <- st_as_sf(svi_2000_geo) %>% st_buffer(0.0001)
# 
# geomac_2000_buff <- st_as_sf(geomac_2000) %>% st_buffer(0.0001)
# 
# joined_buffered_2000 <- 
#   st_join(geomac_2000_buff, svi_2000_buff, left = TRUE, st_overlaps)

## Buffer and join the 2010 data

# svi_2010_buff <- st_as_sf(svi_2010_geo) %>% st_buffer(0.0001)
# 
# geomac_2010_buff <- st_as_sf(geomac_2010) %>% st_buffer(0.0001)
# 
# joined_buffered_2010 <- 
#   st_join(geomac_2010_buff, svi_2010_buff, left = TRUE, st_overlaps)

## Buffer and join the 2014 data

svi_2014_buff <- st_as_sf(svi_2014) %>% st_buffer(0.0001)

geomac_2014_buff <- st_as_sf(geomac_2014) %>% st_buffer(0.0001)

joined_buffered_2014 <- 
  st_join(geomac_2014_buff, svi_2014_buff, left = TRUE, st_overlaps)

## Buffer and join the 2016 data

svi_2016_buff <- st_as_sf(svi_2016) %>% st_buffer(0.0001)

geomac_2016_buff <- st_as_sf(geomac_2016) %>% st_buffer(0.0001)

joined_buffered_2016 <- 
  st_join(geomac_2016_buff, svi_2016_buff, left = TRUE, st_overlaps)

## Buffer and join the 2018 data

svi_2018_buff <- st_as_sf(svi_2018) %>% st_buffer(0.0001)

geomac_2018_buff <- st_as_sf(geomac_2018) %>% st_buffer(0.0001)

joined_buffered <- 
  st_join(geomac_2018_buff, svi_2018_buff, left = TRUE, st_overlaps)

# Read in the SIT 209
sit209 <- 
  read_csv(url_sit209)
```

Now that we're finished reading in all of these many, many datasets, and wrangling them, let's add one more: Karen Short's data from [here](https://www.fs.usda.gov/rds/archive/catalog/RDS-2013-0009.4). We'll use her data as a crosswalk between SIT-209 and MTBS so that we can join both to SVI.

```{r}
# Connect to the database
con <-
  DBI::dbConnect(
    drv = RSQLite::SQLite(),
    dbname = here::here("data-raw", "RDS-2013-0009", "Data", "FPA_FOD_20170508.sqlite"),
  )

# Read the Fires table into a dataframe
karen <- DBI::dbReadTable(con, "Fires")

# Disconect the database
DBI::dbDisconnect(con)
```

```{r}
karen %>% count(FIRE_YEAR)
```


```{r}

# THIS IS MESSED UP. NEED TO KEEP WORKING ON IT
# ALTERNATIVE IS ANOTHER LEFT JOIN WITH FIRES

sit209_karen <-
  sit209 %>% 
  mutate(INC_IDENTIFIER = as.character(INC_IDENTIFIER)) %>% 
  left_join(karen, by = c("INC_IDENTIFIER" = "ICS_209_INCIDENT_NUMBER")) %>% 
  drop_na(MTBS_ID) %>% 
  group_by(MTBS_ID) %>% 
  summarize(
    cost = max(EST_IM_COST_TO_DATE, na.rm = TRUE),
    homes_destroyed = max(SFR_DESTROYED, na.rm = TRUE),
    homes_damaged = max (SFR_DAMAGED, na.rm = TRUE),
    homes_threatened = max(SFR_THREATENED_72, na.rm = TRUE),
    acres = max(CURR_INCIDENT_AREA),
    cost_per_acre = cost / acres,
    homes_destroyed_per_acre = homes_destroyed / acres,
    homes_damaged_per_acre = homes_damaged / acres,
    homes_threatened_per_acre = homes_threatened / acres
  )

ids <- sit209_karen %>% arrange(MTBS_ID) %>% pull(MTBS_ID)

sit209_geo <-
  fires %>% 
  arrange(Fire_ID) %>% 
  filter(Fire_ID %in% ids) %>% 
  bind_cols(sit209_karen)
  
sit_svi_2014 <-
  sit209_geo %>%
  filter(Year == 2014) %>%
  st_intersection(svi_2014)

sit_svi_2015 <-
  sit209_geo %>%
  filter(Year == 2015) %>%
  st_intersection(svi_2016)

```



```{r}
sit_svi_2014 %>% 
  filter(
    RPL_THEMES != -999,
    homes_threatened > 0
  ) %>% 
  mutate(
    cost_per_home_threatened = cost / homes_threatened
  ) %>% 
  filter(cost_per_home_threatened < 3e+07) %>% 
  ggplot(aes(y = cost_per_home_threatened, x = RPL_THEMES)) +
  geom_point(aes(color = acres, size = acres)) +
  scale_color_viridis_c() +
  labs(title = "2014")

sit_svi_2015 %>% 
  filter(
    RPL_THEMES != -999,
    homes_threatened > 0
  ) %>% 
  mutate(
    cost_per_home_threatened = cost / homes_threatened
  ) %>% 
  filter(cost_per_home_threatened < 3e+06) %>% 
  ggplot(aes(y = cost_per_home_threatened, x = RPL_THEMES)) +
  geom_point(aes(color = acres, size = acres)) +
  scale_color_viridis_c() +
  labs(title = "2015")

```




```{r}
# Filter the data to just California

svi_2018_CA <-
  svi_2018 %>% 
  filter(STATE == "CALIFORNIA") 

geomac_2018_CA <-
  geomac_2018 %>% 
  filter(str_detect(fire_num, "2018-CA"))
```


```{r fig.asp = 3}
svi_2018_CA %>% 
  filter(RPL_THEMES != -999) %>% 
  ggplot() +
  geom_sf(aes(fill = RPL_THEMES)) +
  geom_sf(data = geomac_2018_CA, color = "white", fill = NA) +
  scale_fill_viridis_c() +
  theme_void()
```

```{r}
geomac_2018_CA %>%
  filter(fire_name %in% c("CAMP")) %>% 
  ggplot() +
  geom_sf(
    data = 
      svi_2018_CA %>% 
      filter(
        COUNTY %in% c("Butte"),
        RPL_THEMES != -999
      ), 
    aes(fill = RPL_THEMES)
  ) +
  geom_sf(color = "white", fill = NA) +
  scale_fill_viridis_c() +
  theme(legend.position = "bottom")
```

```{r}
joined_buffered %>% 
  filter(
    fire_name == "CAMP",
    COUNTY == "Butte"
  ) %>% 
  group_by(LOCATION) %>% 
  summarize(RPL_THEMES) %>% 
  arrange(desc(RPL_THEMES))
        
```

Census tract north of Oroville versus 


```{r}
svi_acreage_2018 <-
  svi_2018 %>% 
  left_join(as_data_frame(joined_buffered), by = c("FIPS")) %>% 
  mutate(
    has_fire = if_else(is.na(fire_name), "No fire", "Fire"),
    rescaled_acres = scales::rescale(acres, to = c(0,1))
  )

svi_acreage_2016 <-
  svi_2016 %>% 
  left_join(as_data_frame(joined_buffered_2016), by = c("FIPS")) %>% 
  mutate(
    has_fire = if_else(is.na(fire_name), "No fire", "Fire"),
    rescaled_acres = scales::rescale(acres, to = c(0,1))
  )

svi_acreage_2014 <-
  svi_2014 %>% 
  left_join(as_data_frame(joined_buffered_2014), by = c("FIPS")) %>% 
  mutate(
    has_fire = if_else(is.na(fire_name), "No fire", "Fire"),
    rescaled_acres = scales::rescale(acres, to = c(0,1))
  )
```

```{r}
svi_acreage_2018 %>% 
  filter(RPL_THEMES.x != -999) %>% 
  ggplot(aes(y = RPL_THEMES.x)) +
  geom_boxplot(aes(group = has_fire, color = has_fire)) 

svi_acreage_2016 %>% 
  filter(RPL_THEMES.x != -999) %>% 
  ggplot(aes(y = RPL_THEMES.x)) +
  geom_boxplot(aes(group = has_fire, color = has_fire)) 

svi_acreage_2014 %>% 
  filter(RPL_THEMES.x != -999) %>% 
  ggplot(aes(y = RPL_THEMES.x)) +
  geom_boxplot(aes(group = has_fire, color = has_fire)) 
```

```{r}
svi_acreage_2018 %>% 
  drop_na(acres) %>%
  filter(RPL_THEMES.x != -999) %>% 
  ggplot(aes(x = RPL_THEMES.x, y = rescaled_acres)) +
  geom_point() +
  labs(title = "2018")

svi_acreage_2016 %>% 
  drop_na(acres) %>%
  filter(RPL_THEMES.x != -999) %>% 
  ggplot(aes(x = RPL_THEMES.x, y = rescaled_acres)) +
  geom_point() +
  labs(title = "2016")

svi_acreage_2014 %>% 
  drop_na(acres) %>%
  filter(RPL_THEMES.x != -999) %>% 
  ggplot(aes(x = RPL_THEMES.x, y = rescaled_acres)) +
  geom_point() +
  labs(title = "2014")
```



```{r}
svi_acreage_2018 %>% 
  filter(rescaled_acres >= 0.5) %>% 
  count(STATE.x, COUNTY.x, LOCATION.x, RPL_THEMES.x) %>% 
  arrange(desc(RPL_THEMES.x))

svi_acreage_2016 %>% 
  filter(rescaled_acres >= 0.5) %>% 
  count(STATE.x, COUNTY.x, LOCATION.x, RPL_THEMES.x) %>% 
  arrange(desc(RPL_THEMES.x))

svi_acreage_2014 %>% 
  filter(rescaled_acres >= 0.5) %>% 
  count(STATE.x, COUNTY.x, LOCATION.x, RPL_THEMES.x) %>% 
  arrange(desc(RPL_THEMES.x))
```

Shasta County - the Carr fire - 2018
https://censusreporter.org/data/map/?table=B19013&geo_ids=05000US06089,140|05000US06089&primary_geo_id=05000US06089

Monterey County - Soberanes Fire - 2016
https://censusreporter.org/data/map/?table=B19013&geo_ids=05000US06053,140|05000US06053&primary_geo_id=05000US06053 

Okanogan Complex Fire - 2014
https://censusreporter.org/data/map/?table=B19013&geo_ids=05000US53047,140|05000US53047&primary_geo_id=05000US53047#column|B19013001,sumlev|140
https://censusreporter.org/data/map/?table=B19013&geo_ids=05000US53047,140|05000US53047&primary_geo_id=05000US53047#column|B19013001,sumlev|140

```{r}
fips <-
  ussf::boundaries() %>% 
  select(STATEFP, STUSPS) %>% 
  mutate(STATEFP = as.double(STATEFP))

sit209_id <-
  sit209 %>%
  filter(!is.na(POO_STATE_CODE)) %>% 
  left_join(fips, by = c("POO_STATE_CODE" = "STATEFP")) %>% 
  mutate(
    year = lubridate::year(DISCOVERY_DATE),
    month = lubridate::month(DISCOVERY_DATE),
    day = lubridate::day(DISCOVERY_DATE),
    id = as.character(str_glue("{year}{month}{day}{STUSPS}"))
  )
```

# TODOS

* Assign each tract an SVI bucket. For example: 0, (0, .1], (.1, .2], ... 1
* Count the number of fires within those buckets
* Add other features for each bucket, for example: average cost, average size, average homes threatened...maybe medians too

```{r}
svi_buckets <-
  svi_2014 %>% 
  filter(RPL_THEMES != -999) %>% 
  mutate(
    bucket =
      case_when(
        RPL_THEMES == 1 ~ 1,
        RPL_THEMES > 0.95 ~ 0.95,
        RPL_THEMES > 0.9 ~ 0.9,
        RPL_THEMES > 0.85 ~ 0.85,
        RPL_THEMES > 0.8 ~ 0.8,
        RPL_THEMES > 0.75 ~ 0.75,
        RPL_THEMES > 0.7 ~ 0.7,
        RPL_THEMES > 0.65 ~ 0.65,
        RPL_THEMES > 0.6 ~ 0.6,
        RPL_THEMES > 0.55 ~ 0.55,
        RPL_THEMES > 0.5 ~ 0.5,
        RPL_THEMES > 0.45 ~ 0.45,
        RPL_THEMES > 0.4 ~ 0.4,
        RPL_THEMES > 0.35 ~ 0.35,
        RPL_THEMES > 0.3 ~ 0.3,
        RPL_THEMES > 0.25 ~ 0.25,
        RPL_THEMES > 0.2 ~ 0.2,
        RPL_THEMES > 0.15 ~ 0.15,
        RPL_THEMES > 0.1 ~ 0.1,
        RPL_THEMES > 0.05 ~ 0.05,
        RPL_THEMES > 0 ~ 0,
        RPL_THEMES == 0 ~ -0.05,
      )
  )

```


```{r}
## THIS IS THE ONE TO ACTUALLY USE

karen_sit_2 <-
  sit209 %>% 
    mutate(INC_IDENTIFIER = as.character(INC_IDENTIFIER)) %>% 
    left_join(karen, by = c("INC_IDENTIFIER" = "ICS_209_INCIDENT_NUMBER")) %>% 
    drop_na(MTBS_ID) %>% 
    arrange(REPORT_TO_DATE) %>% 
    group_by(MTBS_ID) %>% 
    summarize(
      cost = last(EST_IM_COST_TO_DATE),
      homes_destroyed = last(SFR_DESTROYED),
      homes_damaged = last(SFR_DAMAGED),
      homes_threatened = last(SFR_THREATENED_72),
      acres = last(CURR_INCIDENT_AREA),
      cost_per_acre = cost / acres,
      homes_destroyed_per_acre = homes_destroyed / acres,
      homes_damaged_per_acre = homes_damaged / acres,
      homes_threatened_per_acre = homes_threatened / acres
    )

ids <- karen_sit_2 %>% arrange(MTBS_ID) %>% pull(MTBS_ID)

sit209_geo <-
  fires %>% 
  arrange(Fire_ID) %>% 
  filter(Fire_ID %in% ids) %>% 
  bind_cols(karen_sit_2 )
```


```{r}
bucket_take_2 <-
  sit209_geo %>% 
  filter(Year %in% (c(2014, 2015))) %>% 
  st_join(svi_buckets) %>% 
  group_by(bucket) %>% 
  summarize(
    total_acres = sum(acres, na.rm = TRUE),
    total_cost = sum(cost, na.rm = TRUE),
    total_homes_destroyed = sum(homes_destroyed, na.rm = TRUE),
    total_homes_threatened = sum(homes_threatened, na.rm = TRUE),
    total_homes_damaged = sum(homes_damaged, na.rm = TRUE),
    total_fires = n()
  ) %>% 
  mutate(
    cost_per_acre = total_cost / total_acres,
    cost_per_fire = total_cost / total_fires,
    cost_per_home_threatened = total_cost / total_homes_threatened
  )
```


```{r}
summary(bucket_take_2)
```


```{r}
bucket_take_2 %>% 
  ggplot() +
  geom_col(aes(y = cost_per_acre, x = bucket), position = "dodge")
```



```{r}
bucket_take_2 %>% 
  ggplot() +
  geom_col(aes(y = cost_per_fire, x = bucket), position = "dodge")
```



```{r}
bucket_take_2 %>% 
  ggplot() +
  geom_col(aes(y = total_cost, x = bucket), position = "dodge")
```

```{r}
bucket_take_2 %>% 
  ggplot() +
  geom_col(aes(y = total_homes_threatened, x = bucket), position = "dodge")
```


```{r}
bucket_take_2 %>% 
  ggplot() +
  geom_point(aes(x = total_acres, y = total_cost, color = bucket), size = 3) +
  scale_color_viridis_c()
```

* Should we look at county or Census tract? Problems with sampling error.
* Karen Short's data: It never gets updated. It doesn't match with federal numbers because she goes to every state to ask for state fire data. The states also often collect local data. Federal data usually only covers 1) fires on federal land and 2) fires that are so big the feds have to respond to it. There's a lot of noise because of the small fires. 
* Disaster response happens at the county level. 
* Use urban-rural classification system to determine which fires are in populated places
* SYLVIS LAB_ WUI = densely populated places with fires


