---
title: "R Notebook"
output: html_notebook
---

```{r}
# Libraries
library(tidyverse)
library(jsonlite)
library(sf)

# Parameters

## SIT-209
file_sit209 <- here::here("data-raw", "wildfires_deduped.csv")

#===============================================================================

# Code

# Read in the SIT-209 data
sit209 <-
  file_sit209 %>% 
  read_csv(
    col_types = 
      cols(
        .default = col_character(),
        LATITUDE = col_double(),
        LONGITUDE = col_double(),
        ESTIMATED_COST = col_double(),
        REPORT_DATE = col_datetime(format = ""),
        MAX_PERSONNEL = col_double(),
        INCIDENT_AREA = col_double(),
        MAX_SFR_THREAT = col_double(),
        MAX_MFR_THREAT = col_double(),
        MAX_MU_THREAT = col_double(),
        MAX_COMM_THREAT = col_double(),
        MAX_OUTB_THREAT = col_double(),
        MAX_OTHER_THREAT = col_double(),
        SFR_DAMAGED = col_double(),
        MFR_DAMAGED = col_double(),
        MU_DAMAGED = col_double(),
        COMM_DAMAGED = col_double(),
        OTHER_DAMAGED = col_double(),
        SFR_DESTROYED = col_double(),
        MFR_DESTROYED = col_double(),
        MU_DESTROYED = col_double(),
        COMM_DESTROYED = col_double(),
        OUTB_DESTROYED = col_double(),
        OTHER_DESTROYED = col_double(),
        INJURIES = col_double(),
        FATALITIES = col_double(),
        REPORT_ID = col_character(),
        WILDFIRE = col_logical(),
        STATE_FIPS = col_character(),
        DESTROYED = col_logical(),
        THREATENED = col_logical(),
        CLIMATE_REGION = col_character()
      )
  )
```

```{r}
colnames(sit209)
```

```{r}
summary(sit209) 
```

Filter just to wildfires that have destroyed at least one building:

```{r}
sit209_destroyed <-
  sit209 %>% 
  filter(
    WILDFIRE == TRUE,
    DESTROYED == TRUE
  )
```

Explore the extent of single-family home destruction:

```{r}
sit209_destroyed %>% 
  filter(is.na(SFR_DESTROYED))
```

There are still lots of cases in which a building was destroyed, but not homes.

```{r}
sit209_destroyed %>% 
  drop_na(SFR_DESTROYED) %>% 
  ggplot() +
  geom_boxplot(aes(y = SFR_DESTROYED))
```

Are the two super-outliers legit?

```{r}
sit209_destroyed %>% 
  arrange(desc(SFR_DESTROYED))
```

Yup, this looks about right to me.

Zoom in to try to see reasonable categories.


```{r}
top_destroyed <-
  sit209_destroyed %>% 
  top_n(n = 100, wt = SFR_DESTROYED)
```

```{r}
sit209 %>% 
  group_by(COUNTY, STATE) %>% 
  summarize(
    total_homes_destroyed = sum(SFR_DESTROYED, na.rm = TRUE),
    total_homes_damaged = sum(SFR_DAMAGED, na.rm = TRUE),
    total_fires = n()
  )
```

```{r}
sit209 %>% 
  count(is.na(STATE_FIPS))
```

Geocode attempt 1:

```{r}
sit_209_api <-
  sit209 %>%
  drop_na(LATITUDE, LONGITUDE) %>% 
  mutate(
    LATITUDE = if_else(LATITUDE < 0, LATITUDE * -1, LATITUDE),
    LONGITUDE = if_else(LONGITUDE > 0, LONGITUDE * -1, LONGITUDE),
    geo_url = str_c("https://geocoding.geo.census.gov/geocoder/geographies/coordinates?x=", LONGITUDE,"&y=", LATITUDE, "&benchmark=Public_AR_Current&vintage=Current_Current&layer=10&format=json")
  )  

url_to_county <- function(x) {
  fromJSON(x)[["result"]][["geographies"]][[1]][["COUNTY"]]
}

sit_209_geocoded <-
  sit_209_api %>% 
  head(10) %>% 
  mutate(
    county_lat_long = map(geo_url, url_to_county),
    county_lat_long = pluck(county_lat_long, 1)
  ) 
  
```

Geocode attempt 2:

```{r}
counties <- 
  ussf::boundaries(geography = c("county"), projection = c("longlat")) %>% 
  st_transform(crs = 4269)

sit209_sf <-
  sit209 %>% 
  drop_na(LONGITUDE, LATITUDE) %>% 
  filter(SFR_DESTROYED > 0) %>% 
  mutate(
    LATITUDE = if_else(LATITUDE < 0, LATITUDE * -1, LATITUDE),
    LONGITUDE = if_else(LONGITUDE > 0, LONGITUDE * -1, LONGITUDE)
  ) %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4269)

```

```{r}
lat_long_counties <- 
  st_intersection(counties, sit209_sf) %>%
  st_set_geometry(NULL) %>% 
  select(
    GEOID, 
    NAME,
    colnames(sit209 %>% select(-LONGITUDE, -LATITUDE)),
    -COUNTY
  )
```

```{r}
# Places with county names left over

other_counties <-
  sit209 %>% 
  filter(SFR_DESTROYED > 0) %>% 
  filter(is.na(LONGITUDE) | is.na(LATITUDE)) %>% 
  mutate(
    COUNTY = str_to_title(COUNTY),
    STATE_FIPS = map(STATE_FIPS, ~ if_else(length(.) < 2, str_c("0", .),.)),
    STATE_FIPS = pluck(STATE_FIPS, 1)
  ) %>% 
  left_join(counties, by = c("COUNTY" = "NAME", "STATE_FIPS" = "STATEFP")) %>% 
  drop_na(COUNTYFP) %>% 
  select(
    GEOID, 
    NAME = "COUNTY",
    colnames(sit209 %>% select(-LONGITUDE, -LATITUDE))
  )
```


```{r}
joined <-
  bind_rows(other_counties, lat_long_counties)
```





```{r}
sit209_and_spatial <-
  sit209 %>% 
  left_join(sit209_spatial, by = c("INTERNAL_DB_ID"))
```


```{r}

county_list <-
  ussf::boundaries(geography = "county") %>% pull(NAME)

fips_list <-
  ussf::boundaries(geography = "county") %>% pull(GEOID)

sit209_almost_recoded <-
  sit209_and_spatial %>% 
  mutate(
    county_dbl = as.double(COUNTY),
    county_chr = as.character(COUNTY),
    county_chr_valid = if_else(county_chr %in% counties$NAME, county_chr, "Check original"),
    county_fips_constructed = 
      if_else(!is.na(county_dbl), as.character(str_glue("{STATE_FIPS}{county_dbl}")), NA_character_),
    county_recode = 
      case_when(
        is.na(county_chr) & (is.na(LATITUDE) | is.na(LONGITUDE)) & is.na(county_dbl) ~ NA_character_,
        !(is.na(LATITUDE) & is.na(LONGITUDE)) ~ GEOID,
        is.na(county_chr) & (is.na(LATITUDE) | is.na(LONGITUDE)) & !is.na(county_fips_constructed) ~ county_fips_constructed,
        !is.na(county_chr_valid) & (is.na(LATITUDE) | is.na(LONGITUDE)) & is.na(county_dbl) ~ county_chr_valid,
        TRUE ~ NA_character_
      )
  ) # TODO: match every county_chr_valid and state to a GEOID
#remove all 0s from GEOIDs?
```

```{r}
sit209_almost_recoded %>% 
  summarize_all(~ sum(is.na(.)))
```


# THE NEW PLAN:
* Limit only to the top 100 fires in terms of SF home construction
* Find the counties based on Lat/Long (remember to flip them)
* For the remaining: 1) assign a GEOID based on name or 2) try to find them manually

```{r}
top_destroyed_point <-
  top_destroyed %>%
  drop_na(LATITUDE, LONGITUDE) %>% 
  mutate(
    LATITUDE = if_else(LATITUDE < 0, LATITUDE * -1, LATITUDE),
    LONGITUDE = if_else(LONGITUDE > 0, LONGITUDE * -1, LONGITUDE)
  ) %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4269)

counties <- 
  ussf::boundaries(geography = c("county"), projection = c("longlat")) %>% 
  st_transform(crs = 4269)

top_destroyed_counties <- 
  st_join(top_destroyed_point, counties, join = st_within)
```

```{r}
joined %>% 
  count(GEOID, NAME, STATE, sort = TRUE)
```



```{r}
top_summarized <-
  top_destroyed_counties %>% 
  group_by(NAME, STATE) %>%
  summarize(
    total_cost = sum(ESTIMATED_COST, na.rm = TRUE),
    total_acres = sum(INCIDENT_AREA, na.rm = TRUE),
    total_homes_destroyed = sum(SFR_DESTROYED, na.rm = TRUE),
    total_homes_damaged = sum(SFR_DAMAGED, na.rm = TRUE),
    total_fires = n(),
    total_personnel = sum(MAX_PERSONNEL, na.rm = TRUE)
  )
```


```{r}
joined_summarized <-
  joined %>% 
  mutate(
    GEOID = 
      case_when(
        NAME == "Washington" & STATE == "OK" ~ "40147",
        NAME == "Jackson" & STATE == "FL" ~ "12063",
        NAME == "Jackson" & STATE == "NC" ~ "37099",
        NAME == "Marion" & STATE == "MS" ~ "28091"
      )
  ) %>% 
  group_by(NAME, STATE, GEOID) %>%
  summarize(
    total_cost = sum(ESTIMATED_COST, na.rm = TRUE),
    total_acres = sum(INCIDENT_AREA, na.rm = TRUE),
    total_homes_destroyed = sum(SFR_DESTROYED, na.rm = TRUE),
    total_homes_damaged = sum(SFR_DAMAGED, na.rm = TRUE),
    total_fires = n(),
    total_personnel = sum(MAX_PERSONNEL, na.rm = TRUE)
  ) 
```




```{r}
joined_summarized %>% arrange(desc(total_acres))
```


```{r}
joined_summarized %>% 
  filter(total_acres < 3e+06) %>% 
  ggplot() +
  geom_point(aes(x = total_acres, y = total_cost))
```


```{r}
joined_summarized %>% 
  ggplot() +
  geom_histogram(aes(total_personnel))
```

```{r}
joined_summarized %>% 
  arrange(desc(total_personnel))
```


```{r}
joined_summarized %>% 
  ggplot() +
  geom_histogram(aes(total_cost))
```

```{r}
joined_summarized %>% 
  ggplot() +
  geom_histogram(aes(total_homes_destroyed))
```

```{r}
joined_summarized %>% 
  arrange(desc(total_homes_destroyed)) %>% 
  select(NAME, STATE, total_homes_destroyed, total_fires)
```

```{r}
joined_summarized 
```

```{r}
joined_summarized
```

