---
title: Exploring the SIT-209
author: Amy DiPierro
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(sf)

# Parameters

# Monitoring Trends in Burn Severity: Burned Areas Boundary Dataset
# Source: https://www.mtbs.gov/direct-download
# Download date: 2020-04-21
# Geography: United States
# Years: 2006 - 2017
file_fires <- here::here("data-raw", "mtbs_perimeter_data", "mtbs_perims_DD.shp")

# SIT-209: 2014 - 2017
# Source: BigLocalNews
# Download date: 2020-04-22
# Years: 2014 - 2017
url_sit209 <- "https://stacks.stanford.edu/file/druid:xj043rd8767/xj043rd8767_fire_reports_2014-2017.csv"

#===============================================================================

# Code

# Read in the fire shapefile data
fires <-
  read_sf(file_fires)

# Read in the SVI data
sit209 <- 
  read_csv(url_sit209)

# Variables to select
quantity <- tidyselect::vars_select(names(sit209), contains("QUANTITY"))
personnel <- tidyselect::vars_select(names(sit209), contains("PERSONNEL"))
destroyed <- tidyselect::vars_select(names(sit209), contains("DESTROYED"))
damaged <- tidyselect::vars_select(names(sit209), contains("DAMAGED"))
threatened <- tidyselect::vars_select(names(sit209), contains("THREATENED"))

# States
us <-
  ussf::boundaries(
    geography = "state", 
    projection = "longlat",
  ) %>% 
  filter(!NAME %in% c("Alaska", "Hawaii"))

```

# Wrangle SIT-209

Each row in this dataset appears to be an incremental report of a fire, rather than a report about the total resources the fire has required and the destruction it has wrought.

To determine the total cost, therefore, we'll find the maximum number of homes destroyed and the maximum cost for each unique fire identifier.


```{r}
sit209 %>% 
  group_by(INCIDENT_NAME, INC_IDENTIFIER) %>% 
  summarize(
    cost = max(EST_IM_COST_TO_DATE, na.rm = TRUE),
    homes_destroyed = max(SFR_DESTROYED, na.rm = TRUE),
    homes_damaged = max (SFR_DAMAGED, na.rm = TRUE),
    homes_threatened = max(SFR_THREATENED_72, na.rm = TRUE),
    acres = max(CURR_INCIDENT_AREA),
    cost_per_acre = cost / acres,
    homes_destroyed_per_acre = homes_destroyed / acres,
    homes_damaged_per_acre = homes_damaged / acres,
    homes_threatened_per_acre = homes_threatened / acres
  ) %>%
  filter(homes_threatened > 0, cost > 0) %>% 
  ggplot(aes(x = log2(homes_threatened_per_acre), y = log2(cost_per_acre))) +
  geom_hex() +
  geom_smooth() +
  scale_fill_viridis_c()
```

```{r}
sit209 %>% 
  filter(INC_IDENTIFIER == 7293180) %>% 
  arrange(REPORT_FROM_DATE)
```

```{r}
sit209 %>% 
  group_by(INCIDENT_NAME, INC_IDENTIFIER) %>% 
  arrange(REPORT_FROM_DATE) %>% 
  summarize(
    cost = last(EST_IM_COST_TO_DATE)
  ) %>%
  drop_na(cost) %>% 
  head(10)
```

```{r}
sit209 %>% 
  group_by(INCIDENT_NAME, INC_IDENTIFIER) %>% 
  arrange(REPORT_FROM_DATE) %>% 
  summarize(
    cost = max(EST_IM_COST_TO_DATE, na.rm = TRUE)
  ) %>%
  filter(cost > 0) %>%  
  head(10)
```

```{r}
sit209 %>% filter(!is.na(POO_LATITUDE)) %>% distinct(INC_IDENTIFIER)
```


```{r}
sit209 %>% 
  group_by(INCIDENT_NAME, INC_IDENTIFIER) %>% 
  summarize(
    cost = max(EST_IM_COST_TO_DATE, na.rm = TRUE),
    homes_destroyed = max(SFR_DESTROYED, na.rm = TRUE),
    homes_damaged = max(SFR_DAMAGED, na.rm = TRUE),
    homes_threatened = max(SFR_THREATENED_72, na.rm = TRUE),
    acres = max(CURR_INCIDENT_AREA),
    cost_per_acre = cost / acres,
    homes_destroyed_per_acre = homes_destroyed / acres,
    homes_damaged_per_acre = homes_damaged / acres,
    homes_threatened_per_acre = homes_threatened / acres
  ) %>% 
  filter(
    homes_destroyed > 10,
    acres > 0,
    cost > 0
  ) %>% 
  arrange(desc(homes_destroyed_per_acre))
```

```{r}
sit209 %>% 
  filter(INC_IDENTIFIER == 590661)
```

```{r}
sit209 %>% 
  filter(INC_IDENTIFIER %in% c(4327598, 7292820)) %>% 
  count(INCIDENT_NAME, DISCOVERY_DATE)
```



```{r}
sfr_destroyed <-
  sit209 %>% 
  filter(SFR_DESTROYED > 0) %>% 
  group_by(INCIDENT_NAME, DISCOVERY_DATE) %>% 
  mutate(
    homes_destroyed = sum(SFR_DESTROYED, na.rm = TRUE),
    cost = max(EST_IM_COST_TO_DATE, na.rm = TRUE)
  ) %>% 
  select(INCIDENT_NAME, DISCOVERY_DATE, POO_LATITUDE, POO_LONGITUDE, cost, homes_destroyed) %>% 
  distinct(INCIDENT_NAME, DISCOVERY_DATE, POO_LATITUDE, POO_LONGITUDE, cost, homes_destroyed)
```

Here, we convert lat and long to a geometry column for the purpose of the spatial join.

```{r}
sfr_destroyed_point <- st_as_sf(sfr_destroyed, coords = c("POO_LONGITUDE", "POO_LATITUDE"), crs = 4269, agr = "field")
```

Now, let's find the points that are within the fire perimeters.

```{r}
fire_perimeters <-
  sfr_destroyed_point %>% 
  st_intersection(fires)
```

```{r}
sfr_destroyed %>% 
  ggplot(aes(x = homes_destroyed, y = cost)) +
  geom_point()
```

```{r}
sfr_destroyed %>% 
  filter(homes_destroyed > 5000)
```


```{r}
sfr_destroyed %>% 
  ggplot(aes(cost)) +
  geom_histogram()
```



```{r}


common_ids <-
  fire_perimeters %>% 
  arrange(Fire_ID) %>% 
  pull(Fire_ID)

fires %>% 
  filter(Fire_ID %in% common_ids) %>% 
  arrange(Fire_ID)
```


```{r}
fire_perimeters %>% 
  ggplot() +
  geom_sf(data = us) +
  geom_sf() 
  
```

# Misc

```{r}
fips <-
  ussf::boundaries() %>% 
  select(STATEFP, STUSPS) %>% 
  mutate(STATEFP = as.double(STATEFP))

sit209_id <-
  sit209 %>%
  filter(!is.na(POO_STATE_CODE)) %>% 
  left_join(fips, by = c("POO_STATE_CODE" = "STATEFP")) %>% 
  mutate(
    year = lubridate::year(DISCOVERY_DATE),
    month = lubridate::month(DISCOVERY_DATE),
    day = lubridate::day(DISCOVERY_DATE),
    id = as.character(str_glue("{year}{month}{day}{STUSPS}"))
  )
```


```{r}
fires_id <-
  fires %>% 
  mutate(
    state = str_sub(Fire_ID, end = 2),
    id = as.character(str_glue("{Year}{StartMonth}{StartDay}{state}"))
  )
  
```

```{r}
fires_id %>% 
  left_join(sit209_id, by = c("id")) %>% 
  select(STUSPS, POO_CITY, POO_SHORT_LOCATION_DESC, state)
```

```{r}
sit209 %>% 
  count(is.na(POO_STATE_CODE))
```


```{r}
sit209 %>% 
  group_by(INCIDENT_NAME, DISCOVERY_DATE, POO_STATE_CODE) %>% 
  summarize_at(vars(destroyed), max, na.rm = TRUE) %>%
  mutate(
    total_destroyed = COMMERCIAL_DESTROYED + MFR_DESTROYED + OUTBUILDINGS_DESTROYED + MIXED_USE_DESTROYED + SFR_DESTROYED
  ) %>% 
  arrange(desc(SFR_DESTROYED))
```
